R Notebook

### 2017-06-09 manuscript/draft.Rmd

The mixed model (which formed the basis of our observer-corrected "average"
predictions) found that 70% of the variance in richness in the training set
could be explained by differences among sites, and 21% could be explained by
differences among observers. The remaining 9% represents residual variation,
where a given observer might report a different number of species on different
runs of the same transect. We found remarkably little autocorrelation in these
residuals ([[about 0.04, +/- something]]), suggesting richness is close to
stationary on these time scales, once observer differences are accounted for.

The substantial uncertainty associated with differences among observers caused
reliable improvements in both mean-squared-error and deviance when observer
effects were controlled for (Figure 4). This improvement was smallest for the
"Naive" model, as discussed below.

In general, each model's error increased more-or-less linearly with time,
although the slopes and initial error rates varied across models (Figure 3). The
site-level models (i.e. the baselines and the ARIMA methods) tended to do very
well (Figure 3), especially in terms of RMSE and especially on short time
horizons. For most combinations of year and metric, the best-performing model
was one of the two baselines. Usually, the "naive" model had the lowest error
when observer effects were not explicitly accounted for, and the "average" model
had the lowest error when they were. We discuss the reasons for this difference
in the next section.

Our two error metrics produced qualitatively similar results in most cases, with
the notable exception of the stacked SDMs (Figure 3A and 3B). Deviance, unlike
RMSE, depends on the full predictive distribution of a model, including the
model's estimated uncertainty. As noted elsewhere [@harris2015], these models
are generally far too confident, with observed richness values frequently
falling well outside their confidence limits (Figure 3C). As a result, the
stacked SDMs' deviance was dramatically higher than the next-worst model even
though its RMSE was not unusual. On the other hand, the richness-GBM model and
especially the mistnet JSDM had reasonably well-calibrated uncertainty
estimates, even in the presence of uncorrected observer biases (Figure 3C). This
prevented their deviance from rising as quickly as the site-level models when
the observer population changed (Figure 3B).

In general, the models produced confidence intervals that were too narrow,
indicating overconfident predictions. Of these, the random forest-based SDM
stack was the most overconfident, as discussed above. As one might expect,
correcting for observer differences improved the coverage of most models' 95%
intervals. Interestingly, this effect was very small for mistnet, which seems to
indicate that its latent variables were already accounting for observer-based
uncertainty. The "naive" model was the only model whose confidence intervals
were too wide (Figure 3), which can be attributed to the rapid rate at which
these intervals expand as the time lag increases (Figure 1).

@@

>>>>

@@

Here we have made forecasts using yearly timesteps fifty (?) years into the
future. Many ecological forecasts are temporally aggregated to 5-30 year
timesteps and projected up to a century into the future. These are commonly
beyond the career or even lifespan of the researchers, making reasonable model
assessment impossible. Currently assessments at these large scales can be only
be made with a small number of opportune datasets. For examples hindcasts have
been used to asses models of species richness [@alger2009, distler2015] and
distribution [@rappacciuolo2012, @moran-ordonez2017, @araujo2005,
@eskildsen2013] using data aggregated across similar time scales. More
evaluation studies like these are needed but there is a paucity of adequate long
term data available. Short term forecasts will allow for relatively rapid
validation of models by researchers without needing to wait for enough data to
accumulate. This is not to say they should replace large scale long term
forecasts. The relationships of temporal scaling is currently understudied, but
it is possible they are similar to spatial scale dynamics [@mcgill2010,
@levin1992]. For example drivers of species richness differ at different spatial
scales [@rahbek2001, @hurlbert2003]. Thus short term forecasts should be
expected to add value to end users and inform, not replace, long term larger
scaled forecasts.
